\bf A Hierarchical Model for Heterogenous Reliability Field Data

### 2018-05-13 paper/m-lb-m-technometrics.tex
First pass at edits post 5/13 Sunday meeting

\begin{abstract}
When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failures. The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

@@

Often no information is available to identify why a unit failed, even when the failed units are available. Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive. This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. When this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

@@

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.  A common assumption in series system modeling is the time to failure for the different risks are statistically independent.  Thus, the overall reliability of a unit is the product of the reliability of each of the risks.  Then parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as masking.

Previous papers have employed various assumptions and methods to model masked lifetime failure data.  When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.  \cite{chan} labeled the cause of circuit board failures as early (due to defects), unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the causes of failure.  \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.  \cite{ranjan} considered a competing risk model for early and wearout failures as a mixture of Weibull and exponential distributions.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

\section{Model for field data}
\label{sec:GLFP model}

\subsection{Weibull distribution parameterization}
\label{sec:Weibull parameterization}
The Weibull cumulative
 distribution function (cdf) is
\begin{align}
\label{equation:weibull.cdf}
\Pr(T \leq t|\alpha,\beta ) &= F(t|\alpha,\beta)=1-
\exp \left [-\left (\frac{t}{\alpha} \right )^{\beta}
\right ], \quad t > 0,
\end{align}
where $\beta>0$ is the Weibull shape parameter and $\alpha>0$ is a
scale parameter. Because $\log(T)$ has a smallest extreme value
distribution (a member of the location-scale family of
distributions), the Weibull cdf can also be written as
\begin{align*}
\Pr(T \leq t| \mu,\sigma ) &= F(t| \mu,\sigma)= \Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right], \quad t > 0,
\end{align*}
where $\Phi_{SEV}(z)=1-\exp[-\exp(z)]$ is the standard smallest extreme value
distribution cdf and
$\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
and scale parameters for the distribution of $\log(T)$. In addition to showing the Weibull
distribution is a member of the log-location-scale family of
distributions, this step highlights the generalizability of our approach to other distributions 
that fall into this class.

Going forward we use an alternative parameterization
where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are two important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

@@

The parameter $\pi$ represents the proportion of units susceptible to early failure, and hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

@@

\noindent Note that the $p$ quantile of the GLFP model may be found by setting $p=H(t_p)$ and solving numerically for $t_p$.

@@

\subsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced. Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

\section{Motivating example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, accessed January 18, 2018)\nocite{backblaze}. In addition, Backblaze publishes summary statistics of the different drive-models currently operating.  No other analysis or modeling of the failure data is provided, other than simple rates and proportions.  Backblaze does, however, encourage others to further analyze their data.

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or were in operation.  The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for inclusion of drive-models into our analysis.

@@

\subsection{Analysis of a single drive-model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.  The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures until about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).  Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3},\, 7.1\times 10^{-1} \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,\,1.3 \times 10^2 \rangle. $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

\noindent These prior distributions put probability mass on a wide range of values for all model parameters --- much larger than we would expect for a typical Weibull distribution.  Thus, we consider these prior distributions to be relatively uninformative.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.033 & 0.054 & 0.099 \\ 
 $\beta_1 = 1/\sigma_1 $ & 0.47 & 1.13 & 1.72 \\ 
  $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
  $t_{0.5,1}$ & $1.03$  & $2.28$ & $3.99$\\ 
  $t_{0.2,2}$ & $18.0$ & $18.2$ & $18.6$ \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  Quantiles are in thousands of hours.}
\label{table:1}
\end{table}

Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale family of distributions.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).

@@

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{The estimated GLFP for Drive-model 14 plotted on Weibull paper.  The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are shown in red.  The solid line is an adjusted K-M estimate with pointwise 90\% credible bounds in blue.}
  \label{ex1-overlay}
\end{figure}

\section{Hierarchical GLFP model}
\label{sec:Hierarchical GLFP model}
\subsection{Conditional likelihood}

In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.  Let
\begin{equation}
T_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP model is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.  Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.  All times are again in hours.
\subsection{Parameter distributions}

@@

\subsection{Priors}
The complete GLFP model requires prior distributions on the hierarchical parameters.  We suggest weakly informative proper distributions that allow the data to overwhelm the prior and provide information about the distribution of the population parameters.  Prior knowledge about product lifetimes is useful for choosing the mean of the location parameter ($\eta$) distributions.  However, the corresponding standard deviations should be diffuse to allow the location to vary depending on the observed data.  For the scale parameters ($\tau$) we recommend half-Cauchy distributions as suggested by \citet{gelman2014bayesian}.

@@

This set of model specifications is chosen based on the data, interpretation of the model, as well as estimation considerations.  Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.  Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.  Model 4 allows for the probability of early failure as well as the shape and scale parameters for the wearout failure mode to vary across sub-populations.

For all of the nested models, the parameters for the early failure mode are common across sub-populations.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$). Also the failure-time distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.

\subsection{Model selection}
Bayes factors is a classical apporach to compare alternative model specifications.  However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}. Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  The log pointwise predictive density (lpd), is one method to quantify predictive accuracy,

>>>>

\begin{abstract}
When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failures. The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
Furthermore, when the information carried by lifetime data are limited by sample size, censoring, and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

@@

Often no information is available to identify why a unit failed, even when the failed units are available. Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive. This presents a dilemma for analysts: the data may indicate multiple failure modes, contradicting the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown causes of failure in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. When this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

@@

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.  The general name for such a product is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.  A common assumption in series system modeling is that the times to failure for the different risks are statistically independent.  Thus, the overall reliability of a unit is the product of the reliability of each of the risks.  Then parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as masking.

Previous papers have employed various assumptions and methods to model masked lifetime failure data.  When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.  \cite{chan} labeled the cause of circuit board failures as early (due to defects), unknown, or wearout based on the time of observed failures.  Having knowledge of the failure mode for some of the failures helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the causes of failure.  \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.  \cite{ranjan} considered a competing risk model for early and wearout failures as a mixture of Weibull and exponential distributions.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

\section{Model for field data}
\label{sec:GLFP model}
The methods presented in this paper
could employ any of the commonly-used lifetime distributions such as
the Weibull or lognormal (see Chapters 4 and 5 of
\cite{meeker} for others) to describe either the early
(infant mortality) failures or the wearout failures. To simplify the
presentation and because it adequately describes our example data,
we will use the Weibull distribution to describe both the early and
wearout failures. Changing to any other combination of distributions
in the log-location-scale family of distributions is
straightforward

\subsection{Weibull distribution and reparameterization}
\label{sec:Weibull parameterization}
The Weibull cumulative
 distribution function (cdf) is
\begin{align}
\label{equation:weibull.cdf}
\Pr(T \leq t|\alpha,\beta ) &= F(t|\alpha,\beta)=1-
\exp \left [-\left (\frac{t}{\alpha} \right )^{\beta}
\right ], \quad t > 0,
\end{align}
where $\beta>0$ is the Weibull shape parameter and $\alpha>0$ is a
scale parameter. Because $\log(T)$ has a smallest extreme value
distribution (a member of the location-scale family of
distributions), the Weibull cdf can also be written as
\begin{align*}
\Pr(T \leq t| \mu,\sigma ) &= F(t| \mu,\sigma)= \Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right], \quad t > 0,
\end{align*}
where $\Phi_{SEV}(z)=1-\exp[-\exp(z)]$ is the standard smallest extreme value
distribution cdf and
$\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
and scale parameters for the distribution of $\log(T)$. In addition to showing the Weibull
distribution is a member of the log-location-scale family of
distributions, this step highlights the generalizability of our approach to other distributions 
that fall into this class.

Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are three important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

@@

The parameter $\pi$ represents the proportion of units susceptible to early failure, and hence susceptible to both failure modes. Here the cause of failure is not assumed to be known; thus units from the same population are exchangeable.

@@

\noindent Note that the $p$ quantile of the GLFP model can be found by setting $p=H(t_p)$ and solving numerically for $t_p$.

@@

\subsection{Likelihood}
\label{subsec:likehood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced. Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
\begin{equation}
\label{equation:likelihood}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation}

\noindent This form of the likelihood assumes that the lifetimes are reported as exact
failures. Most reliability and warranty databases have failure times
reported to the nearest hour or day, where such a assumption would
be warranted. In some applications, however, field failure data are
summarized by reporting months is service (e.g., the Nevada format
for warranty data). It is straightforward modify the likelihood to
handle interval censored data by replacing $h(t_i;\bm{\theta})$ in
(\ref{equation:likelihood}) with $H(t_{u i};\bm{\theta})-H(t_{\ell i};\bm{\theta})$
where $t_{\ell i}$ and $t_{u i}$ are, respectively, the lower and
upper endpoints of the interval in which unit $i$
failed.

\section{Motivating example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting daily operational data on all of the hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, accessed January 18, 2018)\nocite{backblaze}. In addition, Backblaze publishes summary statistics of the different drive-models currently operating.  No analysis or modeling of the failure data is provided, other than simple rates and proportions.  Backblaze does, however, encourage others to further analyze their data.

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or were in operation.  The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundreds of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for inclusion of a drive-model into our analysis.

@@

\subsection{Analysis of a single drive-model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.  The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures until about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).  Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and a theoretical justification to use the GLFP model in this application.

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle. $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

\noindent These prior distributions put probability mass on a wide range of values for all model parameters --- much larger than we would expect for a typical Weibull distribution.  Thus, these prior distributions are relatively uninformative.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
  \hline
 & Lower & Estimate & Upper \\ 
  \hline
$\pi$ & 0.033 & 0.054 & 0.099 \\ 
 $\beta_1 = 1/\sigma_1 $ & 0.47 & 1.13 & 1.72 \\ 
  $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
  $t_{0.5,1}$ & $1.03$  & $2.28$ & $3.99$\\ 
  $t_{0.2,2}$ & $18.0$ & $18.2$ & $18.6$ \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  The quantiles are in thousands of hours.}
\label{table:1}
\end{table}

Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale family of distributions.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with confidence bands provides a visual assessment of distributional goodness of fit.  We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).

@@

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{The estimated GLFP model for Drive-model 14 plotted on Weibull paper.  The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are indicated by the smooth region.  The solid line is an adjusted K-M estimate with pointwise 90\% credible bounds in blue.}
  \label{ex1-overlay}
\end{figure}

\section{Hierarchical GLFP model}
\label{sec:Hierarchical GLFP model}
In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.

\subsection{Conditional likelihood}
Let
\begin{equation}
T_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the GLFP model for all subpopulations is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.  Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on fixed values of the parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored. All times are in thousands of hours.  Again, this likelihood
is easy to modify for other situations like interval-censoring,
as explained in Section~\ref{subsec:likehood}.

\subsection{Parameter distributions}

@@

\subsection{Prior Distributions}
The complete GLFP model requires prior distributions on the hierarchical parameters.  We suggest weakly informative proper distributions that allow the data to overwhelm the prior distribution and provide information about the distribution of the population parameters.  Prior knowledge about product lifetimes is useful for choosing the mean of the location parameter ($\eta$) distributions.  However, the corresponding standard deviations should be diffuse to allow the location to vary depending on the observed data.  For the scale parameters ($\tau$) we use half-Cauchy distributions as suggested by \citet{gelman2014bayesian}.

@@

This set of model specifications is chosen based on the data, interpretation of the model, as well as estimation considerations.  Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the lifetime distribution.  Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.  Model 4 allows for the probability of early failure as well as the shape and scale parameters for the wearout failure mode to vary across sub-populations.

For all of the nested models, the parameters for the early failure mode are common across sub-populations.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of the $\pi_g$ across the subpopulations.  Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.

\subsection{Model selection}
The use of Bayes factors is a classical apporach to compare alternative model specifications.  However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}. Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd), is one method to quantify predictive accuracy,
